A Hybrid Neutronics Method for Time-Dependent Control Rod Modeling in Molten Salt Reactors

### 2024-05-28 lit.tex
saaf sn method

Whether one couples separate single-physics software together or develops a multiphysics software
under a single numerical solver framework, multiphysics software for modeling \glspl{MSR} must
employ \textit{tight coupling schemes} to couple the neutronics and thermal-hydraulics governing
equations to capture the strong multiphysics interactions in transient \gls{MSR} simulations
accurately without resorting to very small time-steps \cite{aufiero_development_2014}. Tightly
coupled numerical models handle multiphysics interactions by either updating all state
variables simultaneously in one monolithic solve (\textit{full coupling}) or
iteratively updating all state variables (\textit{fixed point iterations})
until the solution converges in every timestep \cite{keyes_multiphysics_2013}.
Full coupling tends to be more computationally expensive because it combines
all physics equations into a single large system of equations to be solved
simultaneously. In contrast, fixed point iterations involve operator splitting to
separate the system of equations into smaller systems based on their associated
physics, solving smaller systems separately, and iteratively updating the state
variables until convergence. Fixed point iterative methods can be less
stable, less accurate, and have poorer
convergence rates since these methods make iterative corrections
without regard to potentially destabilizing modes introduced by the
multiphysics coupling \cite{keyes_multiphysics_2013}. In particular, Lindsay et al.
\cite{lindsay_introduction_2018} demonstrated the non-convergence of a multiphysics \gls{MSR}
model without delayed neutron precursors when using fixed point iterations. While fully coupled
schemes result in large systems of equations to solve, they can still outperform fixed point
iterative methods in some multiphysics problems through superior stability and convergence rates.
Conversely, proven techniques exist for improving the performance of fixed point iteration coupling
schemes for many relevant computational multiphysics research fields, including reactor analysis
\cite{ragusa_consistent_2009}. Fixed point iterative methods also allow for different temporal and
spatial discretization schemes to meet the requisite resolution of each
individual physics component.

@@

High-fidelity neutron transport methods fall under two general categories: stochastic Monte Carlo
methods and deterministic methods. Monte Carlo methods involve simulating a finite number of
neutron histories. Randomly generated numbers determine the outcome of various probabilistic events
such as travel distances between particle collisions, types of collisions, and scattering angles in
each history until a neutron capture event terminates it. Deterministic methods for neutron
transport, such as the discrete ordinates $S_N$ and spherical harmonics $P_N$ methods, solve the
\gls{BTE} for neutron transport with some approximations for handling the angular and energy
dependence. The \gls{BTE} for neutron transport is given as:
%
\begin{align}
  \frac{1}{v(E)} \frac{\partial}{\partial t} \Psi(\vec{r},E,\hat{\Omega},&t) + \hat{\Omega}\cdot
  \nabla\Psi(\vec{r},E,\hat{\Omega},t) + \Sigma_t(\vec{r},E,t)\Psi(\vec{r},E,\hat{\Omega},t) 
  \nonumber \\
  - \int^\infty_0 dE'& \int_{4\pi} d\hat{\Omega}' \Sigma_s(\vec{r},E'\rightarrow E,\hat{\Omega}'
  \rightarrow \hat{\Omega},t) \Psi(\vec{r},E',\hat{\Omega}',t) \nonumber \\
  =& \frac{\chi(E)}{4\pi}
  \int^\infty_0 dE' \int_{4\pi} d\hat{\Omega}' \nu\Sigma_f(\vec{r},E',t) \Psi(\vec{r},E',
  \hat{\Omega}',t)+S(\vec{r},E,\hat{\Omega}',t) \label{eq:bte}
  \shortintertext{where}
  v(E) =& \mbox{ neutron velocity,} \nonumber \\
  \Psi(\vec{r},E,\hat{\Omega},t) =& \mbox{ neutron angular flux,} \nonumber \\
  \vec{r} =& \mbox{ spatial coordinates,} \nonumber \\
  E =& \mbox{ neutron energy,} \nonumber \\
  \hat{\Omega} =& \mbox{ direction of neutron travel,} \nonumber \\
  t =& \mbox{ time,} \nonumber \\
  \Sigma_t(\vec{r},E,t) =& \mbox{ macroscopic total cross section,} \nonumber \\
  \Sigma_s(\vec{r},E'\rightarrow E,\hat{\Omega}'\rightarrow \hat{\Omega},t) =&
  \mbox{ macroscopic scattering cross section,} \nonumber \\
  \chi(E) =& \mbox{ fission neutron spectrum,} \nonumber \\
  \nu =& \mbox{ number of neutrons produced per fission reaction,} \nonumber \\
  \Sigma_f(\vec{r},E',t) =& \mbox{ macroscopic fission cross section,} \nonumber \\
  S(\vec{r},E,\hat{\Omega}',t) =& \mbox{ external neutron source.} \nonumber
\end{align}

The $S_N$ method discretizes the continuous angular directional phase space into a few discrete
angular directions (ordinates). It uses quadrature rules to replace the integrals over
$\hat{\Omega}$ with summations over the ordinates. On the other hand, the $P_N$ method introduces
Legendre polynomial expansions of the angular flux to approximate the angular dependence. Both
methods require higher-order approximations to produce more accurate flux solutions through more
discrete ordinates or higher-order Legendre expansions. Both methods also discretize the
continuous energy dependence into discrete neutron energy groups, which cover non-overlapping,
finite energy ranges across the entire energy spectrum to form multigroup equations as follows:
%
\begin{align}
  \frac{1}{v_g} \frac{\partial}{\partial t} \Psi_g(\vec{r},\hat{\Omega},&t) + \hat{\Omega}\cdot
  \nabla\Psi_g(\vec{r},\hat{\Omega},t) + \Sigma_{t,g}(\vec{r},t)\Psi_g(\vec{r},\hat{\Omega},t)
  \nonumber \\
  - \sum^G_{g'=1}& \int_{4\pi} d\hat{\Omega}' \Sigma_s^{g'\rightarrow g}(\vec{r},\hat{\Omega}'
  \rightarrow \hat{\Omega},t) \Psi_{g'}(\vec{r},\hat{\Omega}',t) \nonumber \\
  =& \frac{\chi_g}{4\pi}
  \sum^G_{g'=1} \int_{4\pi} d\hat{\Omega}' \nu\Sigma_{f,g'}(\vec{r},t) \Psi_{g'}(\vec{r},
  \hat{\Omega}',t)+S_g(\vec{r},\hat{\Omega}',t) \label{eq:mg-bte}
  \shortintertext{where}
  G =& \mbox{ total number of energy groups,} \nonumber \\
  g =& \mbox{ neutron energy group index (in decreasing energy order)} = 1,2,...,G. \nonumber
\end{align}

@@

On the \gls{TH}-modeling side of \gls{MSR} multiphysics, turbulent flow is a complex, yet important
phenomenon to include in our simulation models to obtain accurate predictions of \gls{MSR} behavior
under various operating conditions. Most \gls{MSR} multiphysics studies with turbulence models rely
on the two-equation turbulence models, which provide acceptable estimates of time-averaged flow
characteristics. One study \cite{podila_cfd_2019} showed close agreement in the\gls{MSRE}
temperature distribution with various one-equation, two-equation, and \gls{RSM} turbulence
models despite relatively larger differences observed in turbulence intensities near the walls.
This observation suggests one-equation models may be adequate for \gls{MSR} designs and simulations
with weaker turbulent effects, although other analyses may require even more accurate turbulence
models like \gls{DES} to capture turbulent effects without time averaging adequately.

>>>>

Whether one couples separate single-physics software together or develops a multiphysics software
under a single numerical solver framework, multiphysics software for modeling \glspl{MSR} must
employ \textit{tight coupling schemes} to couple the neutronics and thermal-hydraulics governing
equations to capture the strong multiphysics interactions in transient \gls{MSR} simulations
accurately without resorting to very small time-steps \cite{aufiero_development_2014}. Tightly
coupled numerical models handle multiphysics interactions by either updating all state
variables simultaneously in one monolithic solve (\textit{full coupling}) or
iteratively updating all state variables (\textit{fixed point iterations})
until the solution converges in every timestep \cite{keyes_multiphysics_2013}.
Full coupling tends to be more computationally expensive because it combines
all physics equations into a single large system of equations to be solved
simultaneously. In contrast, fixed point iterations involve operator splitting to
separate the system of equations into smaller systems based on their associated
physics, solving smaller systems separately, and iteratively updating the state
variables until convergence. Fixed point iterative methods can be less
stable, less accurate, and have poorer
convergence rates since these methods make iterative corrections
without regard to potentially destabilizing modes introduced by the
multiphysics coupling \cite{keyes_multiphysics_2013}. In particular, Lindsay et al.
\cite{lindsay_introduction_2018} demonstrated the non-convergence of a multiphysics \gls{MSR}
model without delayed neutron precursors when using fixed point iterations. While fully coupled
schemes result in large systems of equations to solve, they can still outperform fixed point
iterative methods in some multiphysics problems through superior stability and convergence rates.
Conversely, proven techniques exist for improving the performance of fixed point iteration coupling
schemes for many relevant computational multiphysics research fields, including reactor analysis
\cite{ragusa_consistent_2009}. Fixed point iterative methods also allow for different temporal and
spatial discretization schemes to meet the requisite resolution of each
individual physics model.

@@

High-fidelity neutron transport methods fall under two general categories: stochastic Monte Carlo
methods and deterministic methods. Monte Carlo methods involve simulating a finite number of
neutron histories. Randomly generated numbers determine the outcome of various probabilistic events
such as travel distances between particle collisions, types of collisions, and scattering angles in
each history until a neutron capture event terminates it. Deterministic methods for neutron
transport, such as the discrete ordinates $S_N$ and spherical harmonics $P_N$ methods, solve the
\gls{BTE} for neutron transport with some approximations for handling the angular and energy
dependence. The \gls{BTE} for neutron transport is given as:
%
\begin{align}
  \frac{1}{v(E)} \frac{\partial}{\partial t} \Psi(\vec{r},E,\hat{\Omega},&t) + \hat{\Omega}\cdot
  \nabla\Psi(\vec{r},E,\hat{\Omega},t) + \Sigma_t(\vec{r},E,t)\Psi(\vec{r},E,\hat{\Omega},t) 
  \nonumber \\
  - \int^\infty_0 dE'& \int_{4\pi} d\hat{\Omega}' \Sigma_s(\vec{r},E'\rightarrow E,\hat{\Omega}'
  \rightarrow \hat{\Omega},t) \Psi(\vec{r},E',\hat{\Omega}',t) \nonumber \\
  =& \frac{\chi(E)}{4\pi}
  \int^\infty_0 dE' \int_{4\pi} d\hat{\Omega}' \nu\Sigma_f(\vec{r},E',t) \Psi(\vec{r},E',
  \hat{\Omega}',t)+S(\vec{r},E,\hat{\Omega},t) \label{eq:bte}
  \shortintertext{where}
  v(E) =& \mbox{ neutron velocity,} \nonumber \\
  \Psi(\vec{r},E,\hat{\Omega},t) =& \mbox{ neutron angular flux,} \nonumber \\
  \vec{r} =& \mbox{ spatial coordinates,} \nonumber \\
  E =& \mbox{ neutron energy,} \nonumber \\
  \hat{\Omega} =& \mbox{ direction of neutron travel,} \nonumber \\
  t =& \mbox{ time,} \nonumber \\
  \Sigma_t(\vec{r},E,t) =& \mbox{ macroscopic total cross section,} \nonumber \\
  \Sigma_s(\vec{r},E'\rightarrow E,\hat{\Omega}'\rightarrow \hat{\Omega},t) =&
  \mbox{ macroscopic scattering cross section,} \nonumber \\
  \chi(E) =& \mbox{ fission neutron spectrum,} \nonumber \\
  \nu =& \mbox{ number of neutrons produced per fission reaction,} \nonumber \\
  \Sigma_f(\vec{r},E',t) =& \mbox{ macroscopic fission cross section,} \nonumber \\
  S(\vec{r},E,\hat{\Omega}',t) =& \mbox{ external neutron source.} \nonumber
\end{align}

The $S_N$ method discretizes the continuous angular directional phase space into a few discrete
angular directions (ordinates). It uses quadrature rules to replace the integrals over
$\hat{\Omega}$ with summations over the ordinates. On the other hand, the $P_N$ method introduces
Legendre polynomial expansions of the angular flux to approximate the angular dependence. Both
methods require higher-order approximations to produce more accurate flux solutions through more
discrete ordinates or higher-order Legendre expansions. Both methods also discretize the
continuous energy dependence into discrete neutron energy groups, which cover non-overlapping,
finite energy ranges across the entire energy spectrum to form multigroup equations as follows:
%
\begin{align}
  \frac{1}{v_g} \frac{\partial}{\partial t} \Psi_g(\vec{r},\hat{\Omega},&t) + \hat{\Omega}\cdot
  \nabla\Psi_g(\vec{r},\hat{\Omega},t) + \Sigma_{t,g}(\vec{r},t)\Psi_g(\vec{r},\hat{\Omega},t)
  \nonumber \\
  - \sum^G_{g'=1}& \int_{4\pi} d\hat{\Omega}' \Sigma_s^{g'\rightarrow g}(\vec{r},\hat{\Omega}'
  \rightarrow \hat{\Omega},t) \Psi_{g'}(\vec{r},\hat{\Omega}',t) \nonumber \\
  =& \frac{\chi_g}{4\pi}
  \sum^G_{g'=1} \int_{4\pi} d\hat{\Omega}' \nu\Sigma_{f,g'}(\vec{r},t) \Psi_{g'}(\vec{r},
  \hat{\Omega}',t)+S_g(\vec{r},\hat{\Omega},t) \label{eq:mg-bte}
  \shortintertext{where}
  G =& \mbox{ total number of energy groups,} \nonumber \\
  g =& \mbox{ neutron energy group index (in decreasing energy order)} = 1,2,...,G. \nonumber
\end{align}

@@

On the \gls{TH}-modeling side of \gls{MSR} multiphysics, turbulent flow is a complex, yet important
phenomenon to include in our simulation models to obtain accurate predictions of \gls{MSR} behavior
under various operating conditions. Most \gls{MSR} multiphysics studies with turbulence models rely
on the two-equation turbulence models, which provide acceptable estimates of time-averaged flow
characteristics. One study \cite{podila_cfd_2019} showed close agreement in the \gls{MSRE}
temperature distribution with various one-equation, two-equation, and \gls{RSM} turbulence
models despite relatively larger differences observed in turbulence intensities near the walls.
This observation suggests one-equation models may be adequate for \gls{MSR} designs and simulations
with weaker turbulent effects, although other analyses may require even more accurate turbulence
models like \gls{DES} to capture turbulent effects without time averaging adequately.
